---
title: "Blog Post 3"
author: "Team 6"
date: "2025-03-24"
date-modified: "2025-03-24"
draft: FALSE
---
To clean the NYPD arrest dataset, we focused on selecting only the variables relevant to our analysis - race category, gender, age group, borough of arrest, offense category, and police description of the crime. We removed fully empty columns and unnecessary categories such as arrest date, arrest key, and jurisdiction code. We standardized column names and cleaned string variables by trimming whitespace and applying consistent formatting (e.g., title-casing offense descriptions). All categorical variables, such as age group, race, and borough were converted to factors for easier future analysis. We also eliminated any rows with missing values for easier future analysis as well. Here is the code for our cleaned dataset which can be found under the folder dataset:

library(tidyverse)
library(janitor)
library(lubridate)
raw_data <- read_csv("dataset/NYPD_Arrest_Data__Year_to_Date_.csv")
cleaned_data <- raw_data %>%
  clean_names() %>%                                         
  select(where(~ !all(is.na(.)))) %>%                      
  select(                                                   
    perp_race,           
    perp_sex,            
    age_group,           
    arrest_boro,         
    ofns_desc,           
    pd_desc              
  ) %>%
  mutate(
    perp_race = factor(str_trim(perp_race)),                
    perp_sex = factor(str_trim(perp_sex)),
    age_group = factor(str_trim(age_group)),
    arrest_boro = factor(str_trim(arrest_boro)),
    ofns_desc = str_to_title(str_trim(ofns_desc)),          
    pd_desc = str_to_title(str_trim(pd_desc))
  ) %>%
  drop_na()                                                 
write_rds(cleaned_data, "dataset/cleaned_dataset.rds")

Data Analysis:
```{r}
library(tidyverse)
race_offense_table <- cleaned_data %>%
  count(perp_race, ofns_desc, sort = TRUE) %>%
  group_by(perp_race) %>%
  slice_max(n, n = 5) %>%
  arrange(perp_race, desc(n))

print(race_offense_table)
```

To better understand the relationship between race and type of offense, we created a table displaying the top five offenses by race. This summarized table allows us to identify whether certain offenses are disproportionately represented in certain racial groups and enables us to discover  any unusual or unexpected patterns in the dataset.

```{r}
cleaned_data %>%
  count(perp_sex, ofns_desc, sort = TRUE) %>%
  group_by(perp_sex) %>%
  slice_max(n, n = 3) %>%
  ungroup() %>%
  ggplot(aes(x = reorder(ofns_desc, n), y = n)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ perp_sex, scales = "free_y") +
  labs(
    title = "Top 3 Offense Categories according to Gender",
    x = "Offense Category",
    y = "Number of Arrests"
  )
```
The bar charts display the 3 most common offense categories for males and females in the dataset. While both genders share similar top offenses such as, assault-related charges and petit larceny, males are arrested at a higher frequency for each of the 3 offense categories. This visualization allows for an easier comparison of actual counts, pointing to the disparities in number of offenses respective to gender.

```{r}
cleaned_data %>%
  count(perp_race, sort = TRUE)
```

This table looks at the distribution of arrests by race categories. We found that a disproportionate number of arrests (122049) were recorded for individuals who are 'Black', which is way higher than other races. In comparison, categories including American Indian/Alaskan Native and Unknown had fewer than 1,000 arrests each. These patterns could be showing real-world differences, but also allows us to raise further questions when analyzing our dataset.
