---
title: "Blog Post 6"
author: "Team 6"
date: "2025-04-14"
date-modified: "2025-04-14"
draft: FALSE
---
**Final Steps**
As our team approaches the final stages of the project, we’ve assessed our progress to identify areas of the project where more work is needed. We've successfully conducted an indepth analysis and produced several charts/graphs highlighting demographic differences in violent crime victimization in Chicago. Our visuals include clear breakdowns by sex and race, and support a story about inequalities in crime victimization. We've also formulated a preliminary narrative thesis highlighting the significant disparities based on sex and race to form the foundation of the Big Picture article.

The first task we still need to complete is the interactive dashboard component. We have a template and dataset ready, but we still need to finalize our Shiny dashboard to make sure it supports our narrative by allowing users to explore different demographics interactively. Also, we're editing our Big Picture article’s to create an engaging, clear, and informative style similar to popular media. Next steps include finalizing a creative and engaging headline, finishing the explanatory text around each visualization, and making sure the interactive components flow with the narrative.

```{r}
library(tidyverse)
library(sf)
combined_data <- readRDS("dataset/combined_nypd_acs_with_coords_fixed.rds")
nyc_map <- st_read("dataset/nybb.shp")
```

Arrest Count by Borough
```{r}
combined_data %>%
  count(arrest_boro) %>%
  ggplot(aes(x = reorder(arrest_boro, -n), y = n, fill = arrest_boro)) +
  geom_col() +
  labs(title = "Total Arrests by Borough in 2024", x = "Borough", y = "Number of Arrests") +
  theme_minimal()

```

Arrests vs. Unemployment Rate
```{r}
combined_data %>%
  count(arrest_boro) %>%
  left_join(distinct(combined_data, arrest_boro, unemployment_rate), by = "arrest_boro") %>%
  ggplot(aes(x = unemployment_rate, y = n, label = arrest_boro)) +
  geom_point(size = 4, color = "steelblue") +
  geom_text(vjust = -1) +
  labs(title = "Arrests vs Unemployment Rate by Borough",
       x = "Unemployment Rate (%)", y = "Number of Arrests") +
  theme_minimal()

```
Arrests by Race and Sex
```{r}
combined_data %>%
  count(arrest_boro, perp_race) %>%
  ggplot(aes(x = arrest_boro, y = n, fill = perp_race)) +
  geom_col(position = "dodge") +
  labs(title = "Arrests by Borough and Race", x = "Borough", y = "Number of Arrests", fill = "Race") +
  theme_minimal()

combined_data %>%
  count(arrest_boro, perp_sex) %>%
  ggplot(aes(x = arrest_boro, y = n, fill = perp_sex)) +
  geom_col(position = "dodge") +
  labs(title = "Arrests by Borough and Sex", x = "Borough", y = "Number of Arrests", fill = "Sex") +
  theme_minimal()

```

Top Offense Types in High vs Low Unemployment Areas
```{r}
combined_data %>%
  distinct(arrest_boro, unemployment_rate) %>%
  mutate(unemployment_level = ifelse(unemployment_rate >= median(unemployment_rate, na.rm = TRUE),
                                     "High Unemployment", "Low Unemployment")) -> boro_unemp_group

combined_data %>%
  left_join(boro_unemp_group, by = "arrest_boro") %>%
  count(unemployment_level, ofns_desc) %>%
  group_by(unemployment_level) %>%
  mutate(prop = n / sum(n)) %>%
  slice_max(prop, n = 10, with_ties = FALSE) %>%
  ggplot(aes(x = reorder(ofns_desc, prop), y = prop, fill = unemployment_level)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(title = "Top Offense Types in High vs Low Unemployment Boroughs",
       x = "Offense Description", y = "Proportion", fill = "Unemployment Level") +
  theme_minimal()

```

Arrest Map - Colored by Race
```{r}
arrest_points <- combined_data %>%
  filter(!is.na(longitude), !is.na(latitude)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE)

set.seed(42)
plot_data <- arrest_points %>% sample_n(3000)

ggplot() +
  geom_sf(data = nyc_map, fill = "gray95", color = "black") +
  geom_sf(data = plot_data, aes(color = perp_race), size = 1, alpha = 0.6) +
  labs(title = "Arrest Locations in NYC (2024)",
       subtitle = "Each point represents an arrest. Colored by race.",
       color = "Race") +
  theme_minimal()

```
```{r}
library(tidyverse)
library(sf)

combined_data <- readRDS("dataset/combined_nypd_acs_with_coords_fixed.rds")
nyc_map <- st_read("dataset/nybb.shp")

combined_data <- combined_data %>%
  mutate(unemployment_rate = as.numeric(unemployment_rate))

borough_unemp <- combined_data %>%
  group_by(borough) %>%
  summarise(unemployment_rate = mean(unemployment_rate, na.rm = TRUE))

nyc_map <- nyc_map %>%
  mutate(borough = case_when(
    BoroName == "Manhattan" ~ "MANHATTAN",
    BoroName == "Brooklyn" ~ "BROOKLYN",
    BoroName == "Queens" ~ "QUEENS",
    BoroName == "Bronx" ~ "BRONX",
    BoroName == "Staten Island" ~ "STATEN ISLAND"
  )) %>%
  left_join(borough_unemp %>% select(borough, unemployment_rate), by = "borough") %>%
  select(-matches("^unemployment_rate\\..*"))

arrest_points <- combined_data %>%
  filter(!is.na(longitude), !is.na(latitude)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE)

set.seed(42)
plot_data <- arrest_points %>% sample_n(3000)

ggplot() +
  geom_sf(data = nyc_map, aes(fill = unemployment_rate), color = "white") +
  scale_fill_gradient(low = "lightyellow", high = "darkred", na.value = "grey80") +
  geom_sf(data = plot_data, aes(color = perp_race), size = 1, alpha = 0.6) +
  labs(
    title = "NYC Arrest Locations Overlaid on Borough Unemployment Rates (2024)",
    subtitle = "Boroughs shaded by unemployment rate. Points show arrests colored by race.",
    fill = "Unemployment Rate (%)",
    color = "Race"
  ) +
  theme_minimal()


```
